\thetitle

### 2020-08-18 writing/discussion.tex
discussion: note Retic missing casts/checks

An apparent flaw in our transient is the high cost of blame.
Whereas \citet{vss-popl-2017} report an average slowdown of 2.5x and
 worst-case of 5.4x on fully-typed benchmarks in Reticulated,
 our configurations frequently exceed a 10 minute timeout with transient blame.
Unfortunately, our high cost is closer to the truth;
 Reticulated has two advantages that give low overhead in the prior work.
First, Reticulated's type inference often assigns the dynamic type to local
 variables in "fully typed" code.
This flexible type gives a programmer fewer static errors and less-precise
 blame,; but it also makes the benchmarks run faster.
Second, \citet{vss-popl-2017} use relatively small benchmarks.
Four have since been retired from the official Python benchmark suite
 because they are too small, unrealistic, and unstable.
On larger programs, Reticulated suffers from high overhead.
For example, a Reticulated version of our \texttt{sieve} benchmark runs in
 ~40 seconds without blame, and with blame times out after 10 minutes.

>>>>

An apparent flaw in our transient is the high cost of blame.
Whereas \citet{vss-popl-2017} report an average slowdown of 2.5x and
 worst-case of 5.4x on fully-typed benchmarks in Reticulated,
 our configurations frequently exceed a 10 minute timeout with transient blame.
Unfortunately, our high cost is closer to the truth;
 there are at least three broad issues that skew the earlier results.
First, the 2017 implementation of Reticulated fails to insert certain
 soundness checks\footnote{Missing check: \url{https://github.com/mvitousek/reticulated/issues/36}}
 and blame-map updates\footnote{Missing cast: \url{https://github.com/mvitousek/reticulated/issues/43}}
 from the paper.
Second, Reticulated's type checker assigns the dynamic type to local
 variables in the "fully typed" benchmarks---often because the type system
 cannot articulate a more precise type.
Code that ends up with the dynamic type has fewer constraints to check at run-time.
Third, \citet{vss-popl-2017} use relatively small benchmarks.
Four have since been retired from the official Python benchmark suite
 because they are too small, unrealistic, and unstable.\footnote{Release notes: \url{https://pyperformance.readthedocs.io/changelog.html}}
On larger programs, Reticulated suffers from high overhead.
For example, a Reticulated version of our \texttt{sieve} benchmark runs in
 ~40 seconds without blame, and with blame times out after 10 minutes.
